### 2018-06-20 pubs/itp-2018/sections/proof-reconstruction.tex
[ strip ] first finished draft of its description

The process begins assuming that the user posses a \TPTP valid file that encodes
a \CPL conjecture. If this conjecture is a theorem, by means of \Metis, we
obtain a derivation in \TSTP format of a proof for this conjecture. Now, we use
the \Athena tool to generate the \Agda file that corresponds to the proof-term
for such \TSTP derivation. \Athena parses and prunes the proof-tree of the \TSTP
derivation by removing some redundancies and steps that do not contribute in the
proof. After this translation process, we end up with a proof-term  in the \Agda
file of the natural deduction proof. This \Agda file will make calls to theorems
and lemmas from the \Agda libraries, \texttt{agda-prop} (classical propositional
logic) and \texttt{agda-metis} (propositional reasoning for \Metis derivations).

Using \Agda as the proof-checker, we type-check the proof-terms of the \Agda
file previous generated. If this type-checking success, the \TSTP derivation
generated by \Metis will be \emph{correct} module the \Agda correctness and the
proposed formalisation. In such a case, \Agda will output an interface file.
Otherwise, \Agda will report where the type-checking fails, if it is the case,
the failure must be investigated by the user. The possible causes can be due to
many factors: \emph{completeness} issues in our \Agda libraries for the \Metis
reasoning, bugs in the translation by \Athena, bugs in \Agda, printing errors
of the \TSTP derivation by \Metis or soundness bugs in the \Metis
reasoning.

>>>>

The process begins assuming that the user posses a \TPTP valid file that encodes
a \CPL conjecture. If this conjecture is a theorem, by means of \Metis, we
obtain a derivation in \TSTP format of a proof for this conjecture. Now, we use
the \Athena tool to generate the \Agda file that corresponds to the proof-term
for such \TSTP derivation. \Athena parses and prunes the proof-tree of the \TSTP
derivation by removing some redundancies and steps that do not contribute to
prove the conjecture. When the translation process finishes, we end up with a
proof-term in an \Agda file with the natural deduction proof. In this file, we
make calls to theorems and lemmas in the \Agda libraries, \texttt{agda-prop}
(classical propositional logic) and \texttt{agda-metis} (propositional reasoning
for \Metis derivations).

Using \Agda as the proof-checker, we type-check the proof-terms of the \Agda
file previous generated. If this type-checking success, the \TSTP derivation
generated by \Metis will be \emph{correct} module the \Agda correctness and the
proposed formalisation. In such a case, \Agda will output an interface file.
Otherwise, \Agda will report where the type-checking fails, if so, the failure
must be investigated by the user. The possible causes can be due to many
factors: \emph{completeness} issues in our \Agda libraries for the \Metis
reasoning, bugs in the translation by \Athena, bugs in \Agda, printing errors of
the \TSTP derivation by \Metis or soundness bugs in the \Metis reasoning.

### 2018-06-20 pubs/itp-2018/sections/proof-reconstruction/strip.tex
[ strip ] first finished draft of its description

\begin{remark}
\Metis does not make explicit in the \TSTP derivation the way it uses the
subgoals to prove the goal. We show this approach in the correctness of the
\strip inference rule in Theorem~\ref{thm:strip}. To show this theorem, we prove
Lemma~\ref{lem:inv-uh-lem} and  Lemma~\ref{lem:lem-inv-strip}. In the former, we
introduce the auxiliary function $\fuh_1$ to define the $\fuh$ function used to
define the $\fstrip_1$ function in the latter.
\end{remark}

\begin{mainlemma}
  \label{lem:inv-uh-lem}
Let $\text{n} : \Nat$. If $Γ ⊢ \fuh₁~φ~n$ then $Γ ⊢ φ$ where

@@

The second argument in the $\fuh_1$ function is a natural number that constrains
the number of recursive calls. To provide such a number given a formula
$\varphi$, we call the $\fuh_{cm}$ function defined in~\eqref{eq:uh-complexity}.
This function finds what we call the \emph{function complexity} of the function
$\fuh_1$. For the convenience in the following descriptions, we use the
function $\fuh : \Prop → \Prop$ for $\varphi \mapsto \fuh_{1}~φ~(\fuh_{cm}~φ)$.

\begin{equation}
  \label{eq:uh-complexity}
  \begin{aligned}
    &\hspace{.495mm}\fuh_{cm} : \Prop → \Nat\\
    &\begin{array}{llll}
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ⇒ φ₃)) &= \fuh_{cm}~φ₃ + 2\\
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ∧ φ₃)) &= \fmax~(\fuh_{cm}~φ₂)~(\fuh_{cm}~φ₃) + 1\\
    \fuh_{cm}~φ                &= 0.
    \end{array}
  \end{aligned}
\end{equation}

% \begin{align}
%   \label{eq:uh}
%   \begin{split}
%   &\fuh : \Prop → \Prop\\
%   &\fuh~φ~ = \fuh_{1}~φ~(\fuh_{cm}~φ).
%   \end{split}
% \end{align}

% == CONTINUAR DESDE AQUI

The $\fstrip₀$ function defined in~\eqref{eq:strip-zero} yields the
conjunction of subgoals that implies the goal of the problem in the
\Metis \TSTP derivations. Nonetheless, this function is not a
structurally recursive function (see more details in
Appendix~\ref{app:strip-cm}).
Therefore, we present the $\fstrip₁$ function
in~\eqref{eq:strip-fixed} as the structurally recursive version
of the $\fstrip₀$ function by applying the
bounded technique described in
Section~\ref{ssec:structural-recursion}.
We define the $\fstrip₁$ function based on the reading of the \Metis source code.

@@

In a similar way as we define $\fuh_{cm}$
in~\eqref{eq:uh-complexity}, we
define the $\fstrip_{cm}$ function in Appendix~\ref{app:strip-cm} as
the complexity measure for the $\fstrip₀$ function.
We define the \fstrip function as follows in~\ref{eq:strip}.

\begin{align}
  \label{eq:strip}
  \begin{split}
  &\fstrip : \Prop → \Prop\\
  &\fstrip~φ~ = \fstrip_{1}~φ~(\fstrip_{cm}~φ).
  \end{split}
\end{align}

\begin{mainlemma}
\label{lem:lem-inv-strip}
Let $\text{n} : \Nat$ be the complexity measure of the \fstrip function defined
in~\eqref{eq:strip-fixed}.
If $Γ~⊢~\fstrip₁~φ~n$ then $Γ~⊢~φ$.
\end{mainlemma}

The following theorem is convenient to substitute equals by equals in
a theorem. Recall the equality (≡) is symmetric and transitive as well.
We use these properties without any mention.

\begin{mainlemma}
  \label{lem:subst}
  Substitution theorem.
\begin{equation*}
  \label{eq:substitution-theorem}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ φ$}   \AxiomC{$ψ ≡ φ$}
  \RightLabel{\fsubst}
  \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}
\end{mainlemma}

@@

\begin{mainth}
\label{thm:strip}
Let $\text{n} : \Nat$ be the complexity measure of the strip function defined in Appendix~\ref{app:strip-cm}~\eqref{eq:strip-cm}.
Let $s₂$ and $s₃$ be the subgoals of the goal $φ$, that is,
$$\fstrip₁~φ~n~≡~s₂∧~s₃.$$
If $Γ ⊢ s₂$ and $Γ ⊢ s₃$ then $Γ ⊢ φ$.
\end{mainth}
% -------------------------------------------------------------------

Since \Metis proves a conjecture by refutation,
to prove each subgoal, \Metis assumes the negation of it
by using the \negate rule after the \strip inference application
that introduces such a subgoal.

>>>>

\begin{remark}
\Metis does not make explicit in the \TSTP derivation the way it uses the
subgoals to prove the goal. We show this approach in the correctness of the
\strip inference rule in Theorem~\ref{thm:strip}. To show this theorem, we prove
Lemma~\ref{lem:inv-uh-lem} and  Lemma~\ref{lem:lem-inv-strip}. In the former, we
introduce the auxiliary function $\fuh_1$ to define the $\fuh$ function used in
in the latter in the definition of the $\fstrip$ function.
\end{remark}

\begin{mainlemma}
  \label{lem:inv-uh-lem}
Let be $\text{n} : \Nat$. If $Γ ⊢ \fuh₁~φ~n$ then $Γ ⊢ φ$ where

@@

The second argument in the $\fuh_1$ function is a natural number that constrains
the number of recursive calls. To provide such a number given a formula
$\varphi$, we call the $\fuh_{cm}$ function defined in~\eqref{eq:uh-complexity}.
This function finds what we call the \emph{complexity measure} of the function
$\fuh_1$. For the convenience in the following descriptions, we have the
function $\fuh : \Prop → \Prop$ for $\varphi \mapsto \fuh_{1}~φ~(\fuh_{cm}~φ)$.
\begin{equation}
  \label{eq:uh-complexity}
  \begin{aligned}
    &\hspace{.495mm}\fuh_{cm} : \Prop → \Nat\\
    &\begin{array}{llll}
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ⇒ φ₃)) &= \fuh_{cm}~φ₃ + 2\\
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ∧ φ₃)) &= \fmax~(\fuh_{cm}~φ₂)~(\fuh_{cm}~φ₃) + 1\\
    \fuh_{cm}~φ                &= 0.
    \end{array}
  \end{aligned}
\end{equation}
Now, we define the $\fstrip$ function in \eqref{eq:strip} in a similar way as we
did above for the $\fuh$ function. We have a definition of the $\fstrip₁$ function
in~\eqref{eq:strip-fixed} and a complexity measure denoted by $\fstrip_{cm}$
and defined in~\cite{Prieto-Cubides2017a}.
% \vspace*{-5mm}
\begin{align}
  \label{eq:strip}
  \begin{split}
  &\fstrip : \Prop → \Prop\\
  &\fstrip~φ~ = \fstrip₁~φ~(\fstrip_{cm}~φ).
  \end{split}
\end{align}

The $\fstrip$ function yields the conjunction of subgoals that implies the goal
of the problem in the \Metis \TSTP derivations. Its definition comes mainly from
the reading of the \Metis source code and its \TSTP derivations.

@@

\begin{mainlemma}
\label{lem:lem-inv-strip}
Let $n : \Nat$. If $Γ~⊢~\fstrip₁~φ~n$ then $Γ~⊢~φ$.
\end{mainlemma}

The following lemma is convenient to substitute equals by equals in a theorem.
Recall the equality (≡) is symmetric and transitive, and we use these properties
without any mention.

\begin{mainlemma}[\fsubst]
  \label{lem:subst}
  If $Γ ⊢ φ$ and $ψ ≡ φ$ then $Γ ⊢ ψ$.
% \begin{equation*}
%   \label{eq:substitution-theorem}
%   \begin{bprooftree}
%   \AxiomC{$Γ ⊢ φ$}   \AxiomC{$ψ ≡ φ$}
%   \RightLabel{\fsubst}
%   \BinaryInfC{$Γ ⊢ ψ$}
%   \end{bprooftree}
% \end{equation*}
\end{mainlemma}

@@

\begin{mainth}
\label{thm:strip}
Let $s_1$ and $s_2$ be the subgoals of the goal $φ$, that is,
$\fstrip~\varphi≡~s_{1}∧~s_{2}$.
% $\fstrip₁~φ~n~≡~s₂∧~s₃$ where $\text{n} : \Nat$ is the complexity measure $\fstrip_{cm}$.
If $Γ ⊢ s₂$ and $Γ ⊢ s₃$ then $Γ ⊢ φ$.
\end{mainth}
\begin{proof}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$ Γ ⊢ s_1 $}
  \AxiomC{$ Γ ⊢ s_2 $}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ s_1\wedge s_2$}
  \AxiomC{$\fstrip~φ ≡ s_1\wedge s_2 $}
  \RightLabel{By def.~\eqref{eq:strip}}
  \UnaryInfC{$\fstrip₁~φ~(\fstrip_{cm}~\varphi) ≡ s_1\wedge s_2 $}
  \RightLabel{\fsubst}
  \BinaryInfC{$Γ ⊢ \fstrip₁~φ~n$}
  \RightLabel{Lemma~\ref{lem:lem-inv-strip}}
  \UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
\end{equation*}
\end{proof}

% -------------------------------------------------------------------

\begin{remark}

The user of \Metis will probably note another rule in the derivations related
with these subgoals, the \negate rule. Since \Metis proves a conjecture by
refutation, to prove a subgoal, \Metis assumes its negation with the \negate
rule. This negation ocurrs always after the application of the \strip rule which
as we mentioned before it is the rule that introduces each subgoal.

\end{remark}
